---
output: github_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# COVID-19 Border Quarantine Analysis
Interactive R package for exploring COVID-19 quarantine effectiveness.

## Overview

This package provides an interactive Shiny application for exploring COVID-19 border quarantine effectiveness through the lens of outbreak timing distributions and breach pathways. The analysis is based on data from Zenodo (DOI: 10.5281/zenodo.5802433) and research by Zachreson et al. (2022), which developed computational models of quarantine systems coupled with community transmission dynamics.

The package allows users to examine how different vaccination scenarios influence the distribution of times at which major outbreaks might occur if quarantine breaches happen. A key distinction is made between two types of breaches: those originating from infected travellers and those from quarantine workers. Understanding these pathways and their relative risks under different conditions informs quarantine policy design.

## Installation

You can install the development version of covidanalysis from GitHub with:

```r
# install.packages("remotes")
remotes::install_github("ETC5523-2025/assignment-4-packages-and-shiny-apps-Jingwei-z")
```

## Quick Start

```{r message=FALSE, warning=FALSE}
library(covidanalysis)

# Launch the interactive Shiny app
run_app()

# Or explore the data directly
data(ob_timings_summary)
head(ob_timings_summary)
```


## Usage

### Launch the Shiny Application

To launch the interactive application:

```{r example}
library(covidanalysis)
run_app()
```

The application allows you to explore outbreak timing distributions and breach probabilities across different R₀ values, vaccine efficacy levels, and population vaccination coverage. You can visualize how the 50th and 95th percentiles of outbreak timing change with vaccination parameters, and compare the risk profiles of traveller-initiated versus worker-initiated breaches.

### Access the Data

The package includes three datasets from the Zenodo database:

```{r}
library(covidanalysis)
data(ob_timings)
data(ob_timings_summary)
data(ob_probs_travellers)
head(ob_timings)
head(ob_timings_summary)
head(ob_probs_travellers)
```

- `ob_timings`: Raw distribution of breach times from simulation, showing variability in outbreak timing
- `ob_timings_summary`: Statistical summaries (50th and 95th percentiles) of outbreak timing at each parameter combination
- `ob_probs_travellers`: Outbreak probabilities resulting from traveller and worker breach pathways

## Data Source

All data in this package are sourced from the Zenodo database:

Zachreson, C., Shearer, F. M., Price, D. J., Lydeamore, M. J., McVernon, J., McCaw, J., & Geard, N. (2022). COVID-19 in low-tolerance border quarantine systems: Impact of the Delta variant of SARS-CoV-2. *Science Advances*, 8(14), eabm3624.

DOI: [10.1126/sciadv.abm3624](https://doi.org/10.1126/sciadv.abm3624)

Zenodo Database: [10.5281/zenodo.5802433](https://zenodo.org/records/5802433)

## Key Insights

The outbreak timing distributions reveal substantial variability in when breaches might result in major outbreaks. The 50th percentile indicates the median time until outbreak, while the 95th percentile captures the upper range of this distribution. As vaccination coverage increases and vaccine efficacy improves, both percentiles extend, indicating prolonged protection.

The distinction between traveller and worker breach pathways is significant. Worker-initiated breaches occur less frequently due to rigorous testing schedules, but when they do occur, they represent a substantial risk. Traveller breaches depend on successful exit from quarantine with an undetected infection, which can happen despite multiple test occasions.

## Documentation

For complete documentation, visit the [package website](https://ETC5523-2025.github.io/assignment-4-packages-and-shiny-apps-Jingwei-z/).

For detailed analysis and interpretation of results, see the package vignette:

```{r eval=FALSE}
vignette("covidanalysis_vignette", package = "covidanalysis")
```

## License

This package is licensed under the MIT License - see the LICENSE file for details.

## Author

Jingwei Zhao (jzha0527@student.monash.edu)